---
title: "Week2"
author: "Brian Linn"
date: "January 30, 2017"
output: ioslides_presentation
---

##Slide Title
bullet points and subpoints  
`1. first point`<br>
&nbsp;&nbsp;&nbsp;&nbsp;`a) sub-sub-point`  

bulleted list

- bullet1
- bullet2

numbered list 

1. one  
2. two  

a. lettera  
b. letterb  

##Another Slide
on-click bullet points with > followed by -  

>- point1  
>- point2  
>- point3  

plain text  
*italic text* - single asterisk quotes  
**bold text** - double asterick quotes  

hyperlink  
[link](http://google.com)  

code in single quotes  
'r code in single quotes'  
`codey code code`

##Another Slide
code chunk
``` {r}
#code goes here
x <- 10
y <- 12
x + y
```

##Another Slide
hide the code from the output
```{r hideCode, echo = FALSE}
#this code will not show, but the output will
s <- 22
d <- 22
s + d
```

hide input and output
```{r includeNone, include = FALSE}
#input and output are hidden
library(ggplot2)
```

***
Slide without a title



##Another Slide
###Heading1
####Heading2


##Plot Slide  
```{r plot, echo = FALSE, fig.align='center', fig.cap='My Awesome Figure'}
plot(mtcars$wt, mtcars$mpg, xlab = "Weight", ylab = "MPG")
```